---
output: html_document
editor_options: 
  chunk_output_type: inline
---
Wine Quality Reds Exploration by Jesús Sanz Sanz
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.
library(ggplot2)
library(reshape2)
library(ggplot2)
library(dplyr)
library(maps)
library(RColorBrewer)
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wqr <- read.csv('wineQualityReds.csv')
```

This tidy data set contains 1,599 red wines with 11 variables on the chemical 
properties of the wine. At least 3 wine experts rated the quality of each wine, 
providing a rating between 0 (very bad) and 10 (very excellent).

# Univariate Plots Section

The dataset is related to red variant of the Portuguese "Vinho Verde" wine.  
For more details, consult: http://www.vinhoverde.pt/en/ or the reference 
[Cortez et al., 2009].  
Due to privacy and logistic issues, only physicochemical (inputs) and 
sensory (the output) variables are available (e.g. there is no data about 
grape types, wine brand, wine selling price, etc.).

**Atributes (based on physicochemical tests)**:  
   1 - fixed acidity (tartaric acid - g / dm^3)  
   2 - volatile acidity (acetic acid - g / dm^3)  
   3 - citric acid (g / dm^3)  
   4 - residual sugar (g / dm^3)  
   5 - chlorides (sodium chloride - g / dm^3)  
   6 - free sulfur dioxide (mg / dm^3)  
   7 - total sulfur dioxide (mg / dm^3)  
   8 - density (g / cm^3)  
   9 - pH  
   10 - sulphates (potassium sulphate - g / dm3)  
   11 - alcohol (% by volume)  
**Atribute (based on sensory data)**:   
   12 - quality (score between 0 and 10)  

**Description of attributes**:  
1 - fixed acidity: most acids involved with wine or fixed or nonvolatile 
(do not evaporate readily)  
2 - volatile acidity: the amount of acetic acid in wine, which at too high of 
levels can lead to an unpleasant, vinegar taste  
3 - citric acid: found in small quantities, citric acid can add 'freshness' and 
flavor to wines  
4 - residual sugar: the amount of sugar remaining after fermentation stops, 
it's rare to find wines with less than 1 gram/liter and wines with greater 
than 45 grams/liter are considered sweet  
5 - chlorides: the amount of salt in the wine  
6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial 
growth and the oxidation of wine  
7 - total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2 
oncentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine  
8 - density: the density of water is close to that of water depending on the 
percent alcohol and sugar content  
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic)
to 14 (very basic); most wines are between 3-4 on the pH scale  
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas 
(S02) levels, wich acts as an antimicrobial and antioxidant  
11 - alcohol: the percent alcohol content of the wine  
12 - quality (score between 0 and 10)  

```{r echo=FALSE, Univariate_Plots_Atributes}
str(wqr)
```

```{r echo=FALSE, Univariate_Plots_Summary}
summary(wqr)
```

####1. Fixed acidity (tartaric acid - g / dm^3) 

```{r echo=FALSE, Univariate_Plots_FixedAcidity_Graph}
ggplot(aes(x = fixed.acidity), data = wqr) +
  geom_histogram(color = 'black', fill = 'seagreen2', bins = 40)
```

```{r echo=FALSE, Univariate_Plots_FixedAcidity_Summary}
summary(wqr$fixed.acidity)
```

####2. Volatile acidity (acetic acid - g / dm^3)

```{r echo=FALSE, Univariate_Plots_VolatileAcidity_Graph}
ggplot(aes(x = volatile.acidity), data = wqr) +
  geom_histogram(color = 'black', fill = 'yellow2', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_VolatileAcidity_Summary}
summary(wqr$volatile.acidity)
```

We add to the data set the relationship between volatile and fixed acidity 
to analyze it later.

```{r echo=FALSE, Univariate_Plots_RelativeVolatileAcidity_Graph}
wqr$relative.volatile.acidity <- wqr$volatile.acidity / wqr$fixed.acidity

summary(wqr$relative.volatile.acidity)
```

####3. Citric acid (g / dm^3) 

```{r echo=FALSE, Univariate_Plots_CitricAcid_Graph}
ggplot(aes(x = citric.acid), data = wqr) +
  geom_histogram(color = 'black', fill = 'orange2', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_CitricAcid_Graph2}
ggplot(aes(x = citric.acid), data = wqr) +
  geom_histogram(color = 'black', fill = 'orange2', bins = 100)
```

```{r echo=FALSE, Univariate_Plots_CitricAcid_Summary}
summary(wqr$citric.acid)
```

```{r echo=FALSE, Univariate_Plots_CitricAcid_Table}
table(wqr$citric.acid)
```

The tendency of citric acid is to decrease to 0.75 but three anomalies can be 
observed. Many wines with value 0 or practically 0, others many with 0.49 and 
an extreme value of 1 that could be an erroneous value or a wine that looks for 
a very fruity flavor.

We add to the data set the relationship between citric acid and fixed acidity 
to analyze it later.

```{r echo=FALSE, Univariate_Plots_RelativeCitricAcid_Graph}
wqr$relative.citric.acid <- wqr$citric.acid / wqr$fixed.acidity

summary(wqr$relative.citric.acid)
```

####4. Residual sugar (g / dm^3)

```{r echo=FALSE, Univariate_Plots_ResidualSugar_Graph}
ggplot(aes(x = residual.sugar), data = wqr) +
  geom_histogram(color = 'black', fill = 'snow2', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_ResidualSugar_Graph2}
ggplot(aes(x = residual.sugar), data = wqr) +
  geom_histogram(color = 'black', fill = 'snow2', bins = 30) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_ResidualSugar_Summary}
summary(wqr$residual.sugar)
```

Using a logarithmic axis we can observe a standard normal curve

####5. Chlorides (sodium chloride - g / dm^3)  

```{r echo=FALSE, Univariate_Plots_Chlorides_Graph}
ggplot(aes(x = chlorides), data = wqr) +
  geom_histogram(color = 'black', fill = 'brown', bins = 30)
```

```{r echo=FALSE, Univariate_Plots_Chlorides_Graph2}
ggplot(aes(x = chlorides), data = wqr) +
  geom_histogram(color = 'black', fill = 'brown', bins = 50) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_Chlorides_Summary}
summary(wqr$chlorides)
```

```{r echo=FALSE, Univariate_Plots_Chlorides_Head}
head(subset(wqr, wqr$chlorides < 0.02))
```

It seems that the use of salt is very centralized (a low typical deviation) 
but many samples with much higher values are observed. Maybe to counteract 
other flavors? Perhaps sugar?  
There is also some sample with practically no value. Maybe it's a mistake or 
an exceptional wine that seeks this characteristic.

####6. Free sulfur dioxide (mg / dm^3)

```{r echo=FALSE, Univariate_Plots_FreeSulfurDioxide_Graph}
ggplot(aes(x = free.sulfur.dioxide), data = wqr) +
  geom_histogram(color = 'black', fill = 'khaki', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_FreeSulfurDioxide_Graph2}
ggplot(aes(x = free.sulfur.dioxide), data = wqr) +
  geom_histogram(color = 'black', fill = 'khaki', bins = 20) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_FreeSulfurDioxide_Summary}
summary(wqr$free.sulfur.dioxide)
```

###7. Total sulfur dioxide (mg / dm^3)  

```{r echo=FALSE, Univariate_Plots_TotalSulfurDioxide_Graph}
ggplot(aes(x = total.sulfur.dioxide), data = wqr) +
  geom_histogram(color = 'black', fill = 'khaki3', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_TotalSulfurDioxide_Graph2}
ggplot(aes(x = total.sulfur.dioxide), data = wqr) +
  geom_histogram(color = 'black', fill = 'khaki3', bins = 20) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_TotalSulfurDioxide_Summary}
summary(wqr$total.sulfur.dioxide)
```

```{r echo=FALSE, Univariate_Plots_TotalSulfurDioxide_Head}
head(subset(wqr, wqr$total.sulfur.dioxide > 200))
```

Some extreme value can be observed. We might think that maybe they are low 
quality wines but no. They have a value of 7/10

We add to the data set the relationship between free and total sulfur dioxide 
to analyze it later.

```{r echo=FALSE, Univariate_Plots_RelatedFreeSulfurDioxide_Graph}
wqr$relative.sulfur.dioxide <- wqr$free.sulfur.dioxide / wqr$total.sulfur.dioxide

summary(wqr$relative.sulfur.dioxide)
```

####8. Density (g / cm^3)  

```{r echo=FALSE, Univariate_Plots_Density_Graph}
ggplot(aes(x = density), data = wqr) +
  geom_histogram(color = 'black', fill = 'blue', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_Density_Summary}
summary(wqr$density)
```

####9. PH

```{r echo=FALSE, Univariate_Plots_Ph_Graph}
ggplot(aes(x = pH), data = wqr) +
  geom_histogram(color = 'black', fill = 'gray', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_Ph_Summary}
summary(wqr$pH)
```

```{r echo=FALSE, Univariate_Plots_PH_Head}
head(subset(wqr, wqr$pH < 3 | wqr$pH > 4))
```

We could think that wines with a pH out of 3-4 are not very good wines. We 
can observe that they have ratings between 4-6 and curiously there is one with 
very acid pH 2.88 that has a rating of 8!

####10. Sulphates (potassium sulphate - g / dm3)

```{r echo=FALSE, Univariate_Plots_Sulphates_Graph}
ggplot(aes(x = sulphates), data = wqr) +
  geom_histogram(color = 'black', fill = 'tomato', bins = 20)
```

```{r echo=FALSE, Univariate_Plots_Sulphates_Graph2}
ggplot(aes(x = sulphates), data = wqr) +
  geom_histogram(color = 'black', fill = 'tomato', bins = 20) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_Sulphates_Summary}
summary(wqr$sulphates)
```

```{r echo=FALSE, Univariate_Plots_Sulphates_Head}
head(subset(wqr, wqr$sulphates > 1.5))
```

We can observe that wines with many sulfates are very acidic. We will check 
further if there is a relationship between these two variables.

####11. Alcohol (% by volume)

```{r echo=FALSE, Univariate_Plots_Alcohol_Graph}
ggplot(aes(x = alcohol), data = wqr) +
  geom_histogram(color = 'black', fill = 'purple', bins = 30)
```

```{r echo=FALSE, Univariate_Plots_Alcohol_Graph2}
ggplot(aes(x = alcohol), data = wqr) +
  geom_histogram(color = 'black', fill = 'purple', bins = 20) +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plots_Alcohol_Summary}
summary(wqr$alcohol)
```

####12. Quality (score between 0 and 10)

```{r echo=FALSE, Univariate_Plots_Quality_Graph}
ggplot(aes(x = quality), data = wqr) +
  geom_histogram(color = 'black', fill = 'gold', bins = 6)
```

```{r echo=FALSE, Univariate_Plots_Quality_Summary}
summary(wqr$quality)
```

```{r echo=FALSE, Univariate_Plots_Quality_Discrete}
wqr$quality.ranges <- cut(wqr$quality, breaks = c(0, 4, 6, 10), c('Low', 'Medium', 'High'))
summary(wqr$quality.ranges)
```



Most of the ratings are 5-6. Very few wines get extreme values of 3 or 8. 
I would expect to find some 9 or 10. Perhaps the Portuguese wines do not reach 
such a high quality or perhaps they have not entered into this study because 
of their price.

```{r echo=FALSE, Univariate_Plots_Quality_Head}
head(subset(wqr, wqr$quality == 8))
```

It seems that the wines with the best rating have a high degree of alcohol.

```{r echo=FALSE, Univariate_Plots_Quality_Head2}
head(subset(wqr, wqr$quality == 3))
```

At first glance, there is no generality in its attributes that can be related 
to a low score in valuation.

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 wines in the dataset with 12 features:  
   1 - fixed acidity (tartaric acid - g / dm^3)  
   2 - volatile acidity (acetic acid - g / dm^3)  
   3 - citric acid (g / dm^3)  
   4 - residual sugar (g / dm^3)  
   5 - chlorides (sodium chloride - g / dm^3)  
   6 - free sulfur dioxide (mg / dm^3)  
   7 - total sulfur dioxide (mg / dm^3)  
   8 - density (g / cm^3)  
   9 - pH  
   10 - sulphates (potassium sulphate - g / dm3)  
   11 - alcohol (% by volume)  
   12 - quality (score between 0 and 10)  

The first 11 attributes are numerical and the last one (Quality) is categorical.

Observations:  
* Most quality are 5-6.
* Quality of 8 have high valueos of alcohol.
* The median of alcohol is around 10%.
* Almost all wines have pH between 3-4.
* The density of wine is a little lower than water.

### What is/are the main feature(s) of interest in your dataset?
The main features in the data set are pH, alcohol and quality. Although surely 
all the attributes are representative in the flavor of the wine. Wine is a very 
complex product that involves many factors that derive in its quality.

### What other features in the dataset do you think will help support your \
The density, acidity and sugar/chlorides likely contribute to the quality of
the wine. I think extreme values of the flavours break the balance of flavors
on the palate.

### Did you create any new variables from existing variables in the dataset?
I created the following additional variables:  
1. Relative Volatile Acidity respect Fixed Acidity.
2. Relative Citric Acid respect Fixed Acidity.
3. Relative Free Sulfur Dioxide respect Total Sulfur Dioxide.
4. Quality Ranges of Quality to Low, Medium and High values.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Most distribution are normal or right skewed.  
Acid citric distribution appears bimodal with two peaks around 0 and 0.49.

In addition to including additional variables, it has not been necessary to 
perform any additional operations.

TODO REVISAR

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots_GGPairs}
ggpairs(wqr, columns = c(2, 3, 14, 4, 15, 5, 6, 16, 10, 11, 12, 13, 17), axisLabels = 'internal')
```

We can found the following correlations:  
1. Quality and alcohol (0.476)
2. Quality and volatile acidity (-0.391)
3. sulphates and citric acid (0.313)
4. ph and fixed acidity (-0.683)
5. ph and citrix acid (-0.542)

Other relationships that I want to check are:
1. Quality and relative sulfure
2. Quality and citric acid
3. Quality and fixed acidity 
4. Quality and pH
5. Quality with residual sugar
6. Quality and chlorides

####1. Relative Volatile Acidity

```{r echo=FALSE, Bivariate_Plots_relative.volatile.acidity_Correlation}
with(wqr, cor.test(fixed.acidity, volatile.acidity, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_relative.volatile.acidity_Graph}
plot1 <- ggplot(aes(x = fixed.acidity, y = volatile.acidity), data = wqr) +
  geom_point(alpha = 1/3, color = "yellow3") +
  geom_smooth()

plot2 <- ggplot(aes(x = fixed.acidity, y = relative.volatile.acidity), data = wqr) +
  geom_point(alpha = 1/3, color = "yellow3") +
  geom_smooth()

grid.arrange(plot1, plot2, ncol = 2)
```

We can observe a slight tendency to reduce the volatile acidity by increasing 
the fixed.

####2. Relative Citric Acid

```{r echo=FALSE, Bivariate_Plots_relative.citric.acid_Correlation}
with(wqr, cor.test(fixed.acidity, citric.acid, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_relative.citric.acid_Graph}
plot1 <- ggplot(aes(x = fixed.acidity, y = citric.acid), data = wqr) +
  geom_point(alpha = 1/3, color = "orange3") +
  geom_smooth()

plot2 <- ggplot(aes(x = fixed.acidity, y = relative.citric.acid), data = wqr) +
  geom_point(alpha = 1/3, color = "orange3") +
  geom_smooth()

grid.arrange(plot1, plot2, ncol = 2)
```

We can observe a tendency to increse the fixed acidity by increasing 
the citric acid.

We can summary respect the acidity that the citric acid increase the fixed 
acidity and an increase in this reduces the volatility.

####3. Relative Sulfur Dioxide

```{r echo=FALSE, Bivariate_Plots_relative.sulfur.dioxide_Correlation}
with(wqr, cor.test(total.sulfur.dioxide, free.sulfur.dioxide, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_relative.sulfur.dioxide_Graph}
plot1 <- ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = wqr) +
  geom_point(alpha = 1/3, color = "khaki2") +
  geom_smooth()

plot2 <- ggplot(aes(x = total.sulfur.dioxide, y = relative.sulfur.dioxide), data = wqr) +
  geom_point(alpha = 1/3, color = "khaki2") +
  geom_smooth()

grid.arrange(plot1, plot2, ncol = 2)
```

We can observe a tendency to increse the free sulfur dioxide by increasing 
the total sulfur dioxide. The increase dims as the total increases.

####4. Sulphates and Citric Acid

```{r echo=FALSE, Bivariate_Plots_sulphates_citric.acid_Correlation}
with(wqr, cor.test(sulphates, citric.acid, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_sulphates_citric.acid_Graph}
plot1 <- ggplot(aes(x = sulphates, y = citric.acid), data = wqr) +
  geom_point(alpha = 1/3, color = "tomato2") +
  scale_x_continuous(limits = c(quantile(wqr$sulphates, .01), 
                                quantile(wqr$sulphates, .99))) +
  scale_y_continuous(limits = c(quantile(wqr$citric.acid, .01), 
                                quantile(wqr$citric.acid, .99))) +
  geom_smooth()

plot2 <- ggplot(aes(x = sulphates, y = citric.acid), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, ncol = 2)
```

It is observed that there is a slight relationship between the increase of 
sulphates and the increase of citric acid.

####5. pH and Fixed Acidity

```{r echo=FALSE, Bivariate_Plots_pH_fixed.acidity_Correlation}
with(wqr, cor.test(pH, fixed.acidity, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_pH_fixed.acid_Graph}
plot1 <- ggplot(aes(x = pH, y = fixed.acidity), data = wqr) +
  geom_point(alpha = 1/3, color = "seagreen3") +
  scale_x_continuous(limits = c(quantile(wqr$pH, .01), 
                                quantile(wqr$pH, .99))) +
  scale_y_continuous(limits = c(quantile(wqr$fixed.acidity, .01), 
                                quantile(wqr$fixed.acidity, .99))) +
  geom_smooth()

plot2 <- ggplot(aes(x = pH, y = fixed.acidity), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, ncol = 2)
```

As expected, increasing the acidity decreases the pH to more acid.

####6. pH and Citrix Acid

```{r echo=FALSE, Bivariate_Plots_pH_citric.acid_Correlation}
with(wqr, cor.test(pH, citric.acid, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_pH_citric.acid_Graph}
plot1 <- ggplot(aes(x = pH, y = citric.acid), data = wqr) +
  geom_point(alpha = 1/3, color = "orange3") +
  scale_x_continuous(limits = c(quantile(wqr$pH, .01), 
                                quantile(wqr$pH, .99))) +
  scale_y_continuous(limits = c(quantile(wqr$citric.acid, .01), 
                                quantile(wqr$citric.acid, .99))) +
  geom_smooth()

plot2 <- ggplot(aes(x = pH, y = citric.acid), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, ncol = 2)
```

As expected, as in the previous case, increasing citric acid decreases the pH to more acid.

####7. Quality and pH

```{r echo=FALSE, Bivariate_Plots_quality_pH_Correlation}
with(wqr, cor.test(quality, pH, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_pH_Graph}
plot1 <- ggplot(aes(x = quality, y = pH), data = wqr) +
  geom_jitter(alpha = 1/3, color = "gray") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$pH, .01), 
                                quantile(wqr$pH, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = pH), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = pH), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

The quality and the pH are not related (-0.06) although it can be observed in 
the boxplot that the average of the pH of the wines of more quality slightly 
more acid than in the rest.

####8. Quality and Volatile Acidity

```{r echo=FALSE, Bivariate_Plots_quality_volatile.acidity_Correlation}
with(wqr, cor.test(quality, volatile.acidity, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_volatile.acidity_Graph}
plot1 <- ggplot(aes(x = quality, y = volatile.acidity), data = wqr) +
  geom_jitter(alpha = 1/3, color = "yellow3") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$volatile.acidity, .01), 
                                quantile(wqr$volatile.acidity, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = volatile.acidity), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = volatile.acidity), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

As observed with the correlation value (-0.39) the volatile acid under improves 
the quality of the wine.

####9. Quality and Alcohol

```{r echo=FALSE, Bivariate_Plots_quality_alcohol_Correlation}
with(wqr, cor.test(quality, alcohol, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_alcohol_Graph}
plot1 <- ggplot(aes(x = quality, y = alcohol), data = wqr) +
  geom_jitter(alpha = 1/3, color = "purple") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$alcohol, .01), 
                                quantile(wqr$alcohol, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = alcohol), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = alcohol), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

####10. Quality and Fixed Acidity

```{r echo=FALSE, Bivariate_Plots_quality_fixed.acidity_Correlation}
with(wqr, cor.test(quality, fixed.acidity, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_fixed.acidity_Graph}
plot1 <- ggplot(aes(x = quality, y = fixed.acidity), data = wqr) +
  geom_jitter(alpha = 1/3, color = "seagreen3") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$fixed.acidity, .01), 
                                quantile(wqr$fixed.acidity, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = fixed.acidity), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = fixed.acidity), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

The fixed acidity positively influences the quality of the wine, unlike the 
volatilite acid that affects in an inverse way.

Maybe it's because the acidity in the mouth is positive but not in the nose.

####11. Quality and Residual Sugar

```{r echo=FALSE, Bivariate_Plots_quality_residual.sugar_Correlation}
with(wqr, cor.test(quality, residual.sugar, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_residual.sugar_Graph}
plot1 <- ggplot(aes(x = quality, y = residual.sugar), data = wqr) +
  geom_jitter(alpha = 1/3, color = "snow2") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$residual.sugar, .01), 
                                quantile(wqr$residual.sugar, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = residual.sugar), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = residual.sugar), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

From what we observe in the graphs in the correlation, there is no relation 
between quality and sugar.

####12. Quality and Chlorides

```{r echo=FALSE, Bivariate_Plots_quality_chlorides_Correlation}
with(wqr, cor.test(quality, chlorides, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_quality_chlorides_Graph}
plot1 <- ggplot(aes(x = quality, y = chlorides), data = wqr) +
  geom_jitter(alpha = 1/3, color = "brown2") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$chlorides, .01), 
                                quantile(wqr$chlorides, .99))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green')

plot2 <- ggplot(aes(x = quality.ranges, y = chlorides), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

plot3 <- ggplot(aes(x = quality, y = chlorides), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

From what we observe in the graphs and in the correlation there is a slight 
tendency to improve the quality of the wine with the decrease in chlorides

####12. Alcohol and Density

```{r echo=FALSE, Bivariate_Plots_alcohol_density_Correlation}
with(wqr, cor.test(alcohol, density, method="pearson"))
```

```{r echo=FALSE, Bivariate_Plots_alcohol_density_Graph}
plot1 <- ggplot(aes(x = alcohol, y = density), data = wqr) +
  geom_jitter(alpha = 1/3, color = "dark red") +
  scale_x_continuous(limits = c(quantile(wqr$alcohol, .01), 
                                quantile(wqr$alcohol, .99))) +
  scale_y_continuous(limits = c(quantile(wqr$density, .01), 
                                quantile(wqr$density, .99))) +
geom_smooth()

plot2 <- ggplot(aes(x = alcohol, y = density), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

grid.arrange(plot1, plot2, ncol = 2)
```

There is a clear relationship between alcohol and density. Which was expected 
since the alcohol is less dense than water and therefore its increase reduces 
the density of the wine.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
It is observed that the most valued wines are those with more alcohol, 
more acids (low pH and high fixed acidity) and with low volatile acidity.
It is likely that everything that can bring flavor to the wine improves the 
value of it but without exaggerating the smells of it.
There are a strong relationship between quality and alcohol (0.48). 
Surely this is because wines with more maturity tend to have more quality. 
And the more time in maturation, the longer it takes to ferment the sugar 
in alcohol.  
It would have been very interesting if we had in the data set information about 
the maturation time of the wine.

### Did you observe any interesting relationships between the other features \
At the beginning of the analysis, he expected a clear, directly proportional 
relationship between volatile and fixed acid, citric acid and fixed acid, 
free and total SO2, and citric acid and sulfates.
In all cases, this was the case except in the volatile and fixed rate in 
which volatility decreases with the increase in fixed acidity.

### What was the strongest relationship you found?
The strongest relationship found is between alcohol and density (-0.5) followed 
by alcohol and quality (0.48).

# Multivariate Plots Section

####1. Acids relationships

```{r echo=FALSE, Multivariate_Plots_Acids}
ggplot(aes(x = fixed.acidity, y = volatile.acidity, color = citric.acid), data = wqr) +
  geom_point() +
  facet_wrap(~quality.ranges) +
  scale_colour_gradient(low = "blue", high = "green")
```

As already anticipated, fixed acidity is inversely related to volatile acidity 
but directly to citric acid.

####2. SO2 relationships

```{r echo=FALSE, Multivariate_Plots_S02}
ggplot(aes(x = sulphates, y = total.sulfur.dioxide, color = free.sulfur.dioxide), data = wqr) +
  geom_point() +
  geom_smooth() +
  scale_colour_gradient(low = "snow", high = "brown")
```

There is no relationship between sulphates and sulphides, although there is a 
relationship between free and total sulfur, as previously noted.

####3. Quality positive factors

```{r echo=FALSE, Multivariate_Plots_Positive}
ggplot(aes(x = fixed.acidity, y = pH, color = alcohol), data = wqr) +
  geom_point() +
  facet_wrap(~quality.ranges) +
  geom_smooth() +
  scale_colour_gradient(low = "light blue", high = "dark blue")
```

The tendency between acidity and pH in any type of quality is clear, and the 
presence of higher concentrations of alcohol in high quality wines is clear.

####4. Quality negative factors

```{r echo=FALSE, Multivariate_Plots_Negative}
ggplot(aes(x = quality, y = volatile.acidity, color = chlorides), data = wqr) +
  geom_jitter() +
  geom_line(stat = 'summary', fun.y = median, color = 'orange', linetype = 1) +
  geom_line(stat = 'summary', fun.y = mean, color = 'red', linetype = 2) +
  scale_colour_gradient(low = "yellow", high = "green")
```

The tendency to improve wine quality by decreasing volatile acidity and 
chlorides is also clear.

####5. Linear Model

```{r echo=FALSE, Multivariate_Plots_LinearModel}
m1 <- lm(quality ~ alcohol, data = wqr)
m2 <- update(m1, ~ . + pH)
m3 <- update(m2, ~ . + fixed.acidity)
m4 <- update(m3, ~ . + volatile.acidity)
m5 <- update(m4, ~ . + chlorides)
mtable(m1, m2, m3, m4, m5)
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
There is clearly a low presence of citric acid in the samples of low quality 
and a tendency to low volatile acidity and fixed acidity when increasing the 
quality of the wine.

There is no relationship between sulphates and sulphides, although there is a 
relationship between free and total sulfur, as previously noted.

The tendency to improve wine quality by decreasing volatile acidity and 
chlorides is also clear.

### Were there any interesting or surprising interactions between features?
In general I am surprised by the not very strong relationship between the 
attributes of wine and its evaluation. In some cases, it does not seem to 
influence the residual sugar or SO2.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
Yes, bute the linear model do not have the enought precision (0.325) because the 
relationships between the subjective qualite value are not strongs with the 
attributes of the wines.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x = fixed.acidity, y = volatile.acidity, color = citric.acid), data = wqr) +
  geom_point() +
  facet_wrap(~quality.ranges) +
  scale_colour_gradient(low = "blue", high = "green") +
  ggtitle('Acid Relationships') +
  xlab('Fixed Acidity (g / dm^3)') +
  ylab('Volatile Acidity (g / dm^3)') +
  labs(color = 'Citric Acid (g / dm^3)')
```

### Description One
The Fixed acidity is inversely related to volatile acidity but directly to 
citric acid.
There is clearly a low presence of citric acid in the samples of low quality 
and a tendency to low volatile acidity and fixed acidity when increasing the 
quality of the wine.

### Plot Two
```{r echo=FALSE, Plot_Two}
plot1 <- ggplot(aes(x = quality, y = alcohol), data = wqr) +
  geom_jitter(alpha = 1/3, color = "purple") +
  scale_x_continuous() +
  scale_y_continuous(limits = c(quantile(wqr$alcohol, .05), 
                                quantile(wqr$alcohol, .95))) +
  geom_line(stat = 'summary', fun.y = mean, color = 'dark green') +
  ggtitle('Alcohol-Quality Relationships') +
  xlab('Quality (0-10)') +
  ylab('Alcohol (% by Volume)')

plot2 <- ggplot(aes(x = quality.ranges, y = alcohol), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) +
  ggtitle('Alcohol-Quality Relationships') +
  xlab('Quality Ranges') +
  ylab('Alcohol (% by Volume)')

plot3 <- ggplot(aes(x = quality, y = alcohol), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) +
  ggtitle('Alcohol-Quality Relationships') +
  xlab('Quality (0-10)') +
  ylab('Alcohol (% by Volume)')

grid.arrange(plot1, plot2, plot3, ncol = 3)
```

### Description Two
There are a strong relationship between quality and alcohol (0.48). 
Surely this is because wines with more maturity tend to have more quality. 
And the more time in maturation, the longer it takes to ferment the sugar 
in alcohol.  
It would have been very interesting if we had in the data set information about 
the maturation time of the wine.

### Plot Three
```{r echo=FALSE, Plot_Three}
plot1 <- ggplot(aes(x = alcohol, y = density), data = wqr) +
  geom_jitter(alpha = 1/3, color = "dark red") +
  scale_x_continuous(limits = c(quantile(wqr$alcohol, .05), 
                                quantile(wqr$alcohol, .95))) +
  scale_y_continuous(limits = c(quantile(wqr$density, .05), 
                                quantile(wqr$density, .95))) +
geom_smooth() +
  ggtitle('Alcohol-Density Relationships') +
  xlab('Alcohol (% by Volume)') +
  ylab('Density (g / cm^3)')

plot2 <- ggplot(aes(x = alcohol, y = density), data = wqr) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) +
  ggtitle('Alcohol-Density Relationships') +
  xlab('Alcohol (% by Volume)') +
  ylab('Density (g / cm^3)')

grid.arrange(plot1, plot2, ncol = 2)
```

### Description Three
The most clear relationship in this dataset is between alcohol and density 
(-0.5). Which was expected since the alcohol is less dense than water and 
therefore its increase reduces the density of the wine.

------

# Reflection
In conclusion we can say that the factor that contributes most to the quality 
of the wine (subjective evaluation of the jury) is the time of maturation of 
the wine. The time is not among the attributes of the dataset but it is deduced 
by the amount of wine alcohol that is the result of the maturation process 
which requires a lot of time.  
Other factors that positively influence the quality ofthe wine are all the 
nuances in the taste and smell that come from the balance of acids found in it.  
The acid aroma in the aroma of the wine decreases the value of it as observed 
in the study.

The biggest difficulty I have found in this dataset is to find clear 
relationships between the attributes. A quality wine is full of flavors, 
nuances, textures and smells. And each of these characteristics will be 
reflected differently in the chemical attributes of the wine.  
It is clear that it is complicated to be an enologist!

In part, the not very strong relationship between attributes and valuation have 
led to the fact that the linear prediction model is obviously not accurate.

It would have been very interesting to have data such as the maturity of the 
wine, the type of process it has carried, the type of grape, type of barrel 
that has been used, etc. The winemaking process is very complex and any factor 
can vary the final result.